---
title: "blast_to_RFDB"
author: "Erin D'Agnese"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script is used to build a reference DB for COI based off Blast results. 
The goal of this script is to create two files:
1: a fasta file with an identifier/accession number and the representative sequence that has been trimmed to the COI region
2: a text file with the accession/identifier and the taxonomic assignment for the rep sequence in the db

1. Fist step is running this command through the server to blast the results and pull down the data we need for each
```{bash}
#!/bin/bash

BLAST_DB='///_Ev.db'
# BLAST PARAMETERS
PERCENT_IDENTITY="85"
WORD_SIZE="30"
EVALUE="1e-30"
# number of matches recorded in the alignment:
MAXIMUM_MATCHES="50"
CULLING="5"

	################################################################################
	# BLAST CLUSTERS
	################################################################################
	echo $(date +%H:%M) "BLASTing..."
	blast_output="//coi_blast_20210723.txt"
blastn \
		-query "////hash_key.fasta" \
		-db "${BLAST_DB}" \
		-num_threads 4 \
		-perc_identity "${PERCENT_IDENTITY}" \
		-word_size "${WORD_SIZE}" \
		-evalue "${EVALUE}" \
		-max_target_seqs "${MAXIMUM_MATCHES}" \
		-culling_limit="${CULLING}" \
		-outfmt "6 qseqid sseqid sacc pident length mismatch gapopen qcovus qstart qend sstart send evalue bitscore staxids qlen sscinames sseq" \
		-out "${blast_output}"
```

```{r}
library(insect)
library(tidyverse)
library(taxonomizr)
```


2. Using the results from Blast we add headers to the dataframe and convert taxaid to numeric where possible, then find the ones that coerce NAs because they have more than one taxID
```{r}
blast <- read_delim(("/Users/erdag/NextGenNEPA_local/classifiers/COI_build/runs1_2_3_coi_hash_key_format6-1.txt"), col_names = c("qseqid", "sseqid", "sacc", "pident", "length", "mismatch", "gapopen", "qcovus", "qstart", "qend", "sstart", "send", "evalue", "bitscore", "staxid", "qlen", "sscinames", "sseq"), delim = "\t" )

blast %>% mutate (Ntaxid = as.numeric (staxid)) %>% filter(is.na(Ntaxid))
```

There are a number of hashes that have more than 1 taxID which are not being properly handled. To deal with this, we can try a couple things, first could try creating a table of just these ones that we can run separately after making decisions on which TaxID to take. 

```{r}
multi_tax <- blast %>% mutate (Ntaxid = as.numeric (staxid)) %>% filter(is.na(Ntaxid))
#there are 888 entries that have multiple taxIDs per hash
nblast<- blast %>% mutate (taxid = as.numeric (staxid))
```



3.pull the taxonomy using insect 
```{r}
insect::taxonomy() -> worlds.taxonomy
```

4. combine the taxonomy with the blast results
```{r}
blast %>% distinct(staxid) %>%
  pull()-> all.taxids
all.taxids <- map(all.taxids, as.numeric)
insect::get_lineage(all.taxids, worlds.taxonomy) -> all.taxonomy   
```



5. make a dataframe with all the ranks for each taxID
```{r}
Conversion.df <- tibble (taxid = all.taxids,
                         taxonomy = all.taxonomy) %>%
  mutate (Kingdom = map_chr(taxonomy, ~.x["kingdom"]),
          Phylum = map_chr(taxonomy, ~.x["phylum"]),
          Class = map_chr(taxonomy, ~.x["class"]),
          Order = map_chr(taxonomy, ~.x["order"]),
          Family= map_chr(taxonomy, ~.x["family"]),
          Genus = map_chr(taxonomy, ~.x["genus"]),
          Species = map_chr(taxonomy, ~.x["species"]))

Conversion.df %>% filter(is.na(taxid))

```

6. merge the Conversion.df with the blast df to run the consensus script
```{r}
blast.tax <- merge(nblast, Conversion.df, by= "taxid")
```

7. Pull out the accession and taxonomy for the text file
```{r}
ref.txt <- blast.tax %>% select(sacc, Kingdom, Phylum, Class, Order, Family, Genus, Species)
ref.txt <- ref.txt %>% unite(tax, c(Kingdom, Phylum, Class, Order, Family, Genus, Species), sep = ";")
write.table(ref.txt, "/Users/erdag/NextGenNEPA_local/classifiers/COI_build/COI_blastDB.txt", sep="\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
```

8. Pull out the accession and sequence for the fasta
```{r}
ref.fasta <- blast.tax %>% select(sacc, sseq)
names(ref.fasta)[1]<-'name'
names(ref.fasta)[2]<-'seq'

#remove duplicates
nref.fasta <- ref.fasta %>% distinct(name, .keep_all = TRUE)

writeFasta<-function(data, filename){
  fastaLines = c()
  for (rowNum in 1:nrow(data)){
    fastaLines = c(fastaLines, as.character(paste(">", data[rowNum,"name"], sep = "")))
    fastaLines = c(fastaLines,as.character(data[rowNum,"seq"]))
  }
  fileConn<-file(filename)
  writeLines(fastaLines, fileConn)
  close(fileConn)
}


writeFasta(ref.fasta, "/Users/erdag/NextGenNEPA_local/classifiers/COI_build/COI_blastRefDB.fasta")

```



Now we need to deal with the multi-taxids
```{r}
#create df with accession numbers and the taxids from multi_tax
multi_tax_ids <- multi_tax %>% select(sacc, staxid)
#remove duplicate rows by accession numbers
multi_tax_ids <- multi_tax_ids %>% distinct(sacc, .keep_all = TRUE)
#splitting the multiple vlaues in staxid to create multiple rows
library(tidyr)
multi_tax_sep <- separate_rows(multi_tax_ids, staxid,sep=";")
#making a df with the taxonomic ranks 
ntaxid<- multi_tax_sep %>% mutate (taxid = as.numeric (staxid))
multi_tax_sep %>% distinct(staxid) %>%
  pull()-> multi.taxids
multi.taxids <- map(multi.taxids, as.numeric)
insect::get_lineage(multi.taxids, worlds.taxonomy) -> multi.taxonomy   
#convert to df with taxonomy
multi.df <- tibble (taxid = multi.taxids,
                         taxonomy = multi.taxonomy) %>%
  mutate (Kingdom = map_chr(taxonomy, ~.x["kingdom"]),
          Phylum = map_chr(taxonomy, ~.x["phylum"]),
          Class = map_chr(taxonomy, ~.x["class"]),
          Order = map_chr(taxonomy, ~.x["order"]),
          Family= map_chr(taxonomy, ~.x["family"]),
          Genus = map_chr(taxonomy, ~.x["genus"]),
          Species = map_chr(taxonomy, ~.x["species"]))

multi.df %>% filter(is.na(taxid))

```

9. Create the function to find consensus in the accession numbers of the taxonomic ranks associated with the multiple taxids
```{r}
custom2.lca <- function (df) {df %>%  # this function allows to change cutoff parameters for a specified dataframe (df)
  group_by(sacc) %>%
  select(Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  nest() %>% # for each query, calculate the agreed taxonomy
  # ungroup %>% slice (1:10) %>%
  mutate(consensus = purrr::map(data,  function(.x) { 
    # If there are 100% matches - keep those and the 90s
    # If not, keep everything
.x %>% 
  condenseTaxa() %>%
      paste(., collapse = ";")}

                               )) %>%
  select(sacc, consensus) %>%
  unnest(consensus)} 
```



8.make a dataframe with the consensus for each hash
```{r}

#merge the accession numbers with the new df
multi.tax <- merge(ntaxid, multi.df, by= "taxid")

consensus.multi <- custom2.lca(multi.tax)

#rename the the tax column to match ref.txt
names(consensus.multi)[2]<-'tax'

##pull out the seqs and accession numbers into a df to merge
multi.fasta <- multi_tax %>% select(sacc, sseq)
#remove the duplicates and rename the columns
multi.fasta <- multi.fasta %>% distinct(sacc, .keep_all = TRUE)
names(multi.fasta)[1]<-'name'
names(multi.fasta)[2]<-'seq'

```

now merge the taxonomy files and create a text file with all of them in there
```{r}
coi_DB_taxonomy <- rbind(ref.txt, consensus.multi)
#remove the duplicates so there is one row for each accession number
coi_DB_taxonomy <- coi_DB_taxonomy %>% distinct(sacc, .keep_all = TRUE)

write.table(coi_DB_taxonomy, "/Users/erdag/NextGenNEPA_local/classifiers/COI_build/COI_blast_RefDB.txt", sep="\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

```

merge the fasta dataframes and create a fasta file with all of them in there
```{r}
coi_DB_fasta <- rbind(ref.fasta, multi.fasta)
#remove the duplicates so there is one row for each accession number
coi_DB_fasta <- coi_DB_fasta %>% distinct(name, .keep_all = TRUE)


writeFasta(coi_DB_fasta, "/Users/erdag/NextGenNEPA_local/classifiers/COI_build/COI_blast_RefDB.fasta")
coi_DB_fasta <- readDNAStringSet("/Users/erdag/NextGenNEPA_local/classifiers/COI_build/COI_blast_RefDB.fasta")

name = names(coi_DB_fasta)
seq = paste(coi_DB_fasta)
coi_DB_fasta <- data.frame(name, seq)

coi_DB_taxonomy <- read.table("/Users/erdag/NextGenNEPA_local/classifiers/COI_build/COI_blast_RefDB.txt", sep = "\t", header = FALSE)
names(coi_DB_taxonomy)[1] <- 'name'
names(coi_DB_taxonomy)[2] <- 'tax'

#make a df combining our blast dfs for use in combining with wangensteen later
names(coi_DB_taxonomy)[1]<-'name'
blast.all <- coi_DB_taxonomy %>% left_join(coi_DB_fasta, by=c("name" = "name"))
```

read in the wangensteen fasta
```{r}
library("Biostrings")
wang.fasta <- readDNAStringSet("/Users/erdag/NextGenNEPA_local/classifiers/COI_build/Wangensteen_RF_DB.fasta")

name = names(wang.fasta)
seq = paste(wang.fasta)
wang.fasta.df <- data.frame(name, seq)
```

read in the text and format to merge
```{r}
wang.tax <- read.table("/Users/erdag/NextGenNEPA_local/classifiers/COI_build/Wangensteen_RF_DB_taxonomy.txt", sep = ",", header = TRUE)

wang.tax.df <- wang.tax %>% unite(tax, c(kingdom_names, phylum_names, class_names, order_names, family_names, genus_names, species_names), sep = ";")
names(wang.tax.df)[1]<-'name'

##make df with accession, tax, and seq
wang.all <- wang.tax.df %>% left_join(wang.fasta.df, by=c("name" = "name"))

```

creating the merged taxonomy and text files
```{r}
#first find the sequences that are only in Wangesteen's DB so we only add those to the ones we've blasted
new.wang <- anti_join(wang.all, blast.all, by=c("seq"="seq"))

#combine our blast results and tax with the ones from Wangensteen that weren't duplicates of ours
blast.new.wang <- rbind(blast.all, new.wang)

# create the new final txt file
ref.db.tax <- blast.new.wang %>% select(name,tax)
write.table(ref.db.tax, "/Users/erdag/NextGenNEPA_local/classifiers/COI_build/Combined_blast_wang_RefDB_taxonomy.txt", sep="\t", row.names = FALSE, col.names = FALSE, quote = FALSE)

#create new final fasta file
ref.db.fasta <- blast.new.wang %>% select(name, seq)

writeFasta(ref.db.fasta, "/Users/erdag/NextGenNEPA_local/classifiers/COI_build/Combined_blast_wang_RefDB.fasta")


```


