---
title: "hashes_to_taxonomy"
author: "Erin D'Agnese"
date: "6/1/2021"
output:   
  html_notebook:
    toc: true
params: 
  folder:
    value: C:\Users\erdag\github\NextGenNEPA\Output\hashes_to_taxonomy_output\run4_preliminary\COI 
  Hash_key:
    value: C:\Users\erdag\github\NextGenNEPA\Output\dada2_output_files\run4_preliminary\COI_hash_key.csv
  ASVs: 
    value: C:\Users\erdag\github\NextGenNEPA\Output\dada2_output_files\run4_preliminary\COI_ASV_table.csv
  classifier:
    value: C:\Users\erdag\NextGenNEPA_local\classifiers\classification.tree.coi.leray.v5.rds
  previous_effort: 
    value: C:\Users\erdag\github\NextGenNEPA\Output\hashes_to_taxonomy_output\run3\COI\2021-10-18hashes.annotated.rds
 
---

Classifying the Dada2 output sequences using the classifier created using insect from the CRUX reference databases
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = params$folder)
```


```{r load libraries, echo = FALSE}
library (tidyverse)
library (insect)
library (seqinr)
library (googlesheets)
library (here)
library (patchwork)
```


```{r load objects by the end of the cleaning }
Hash     <- read_csv(params$Hash_key) %>% 
  select(Hash, Sequence) %>% distinct()
ALL.ASVs <- read_csv(params$ASVs)
tree.2   <- read_rds(params$classifier)

#bringing in the previous annotated run to start from and add to
previous.effort <- read_rds(params$previous_effort)
```

So these sequences are in the same direction, and they have some shared and some new sequences. 

Let's keep only the unique sequences


Make them into a DNAbin object for insect

```{r make it into a DNA object}
new.set <- anti_join(Hash, previous.effort, by = c("Hash" = "representative"))
all.hashes.insect <- char2dna(new.set$Sequence)
names (all.hashes.insect) <- new.set$Hash
all.hashes.insect
```


classify all sequences in our bin file

```{r classify}
clasif.hashes <- classify (x = all.hashes.insect, tree = tree.2, cores = 4)

clasif.hashes %>% 
  unite (family, genus, species, sep = "|", col = "taxa")
clasif.hashes %>% dplyr::count (rank) %>% arrange(desc(n))
```


OK. So now let's save the classification object as an RDS and CSV

```{r save it for now}
clasif.hashes %>% 
  filter(family!= "" & phylum == "") %>% 
  distinct(class) # How many have a valid family but no phylum info
# Add new phylum info
clasif.hashes %>% 
  mutate(phylum = case_when(phylum != "" ~ phylum,
                            TRUE   ~ class))

saveRDS(clasif.hashes, paste0("hashes.annotated", ".rds"))
clasif.hashes <- readRDS("hashes.annotated.rds")
write.csv(clasif.hashes, "hashes.annotated.csv")

```

apply thresholds and save modified .rds

```{r apply thresholds}
thresholds <- list(0.8, 0.85, 0.95)
thresholds.classif <- map(thresholds, ~ classify(x= all.hashes.insect,
                                              tree = tree.2,
                                              cores = 8,
                                              threshold = .))
names(thresholds.classif) <- thresholds
#ADD THE DEFAULT
thresholds.classif[[4]] <- clasif.hashes
# Name it
names(thresholds.classif)[4] <- "0.9"
saveRDS(thresholds.classif, file =paste0("hashes.annotated.threshold.rds"))
list.of.thres <- readRDS(file ="hashes.annotated.threshold.rds")
l2 <- lapply (list.of.thres, function (x) as_tibble(x))

```

Check the classification: the resolution of dataset, and classifier, let's see how many reads can be classified to species, genus, family and order level

```{r checking the classification}
clasif.hashes %>% dplyr::count (rank) %>% arrange(desc(n))# an overview of the taxonomic resolution
# a first check: How many sequences did not have a score: Either bc they exactly match a sequence in the tree
clasif.hashes %>% 
  filter (is.na(score)) %>% # 176 Hashes
  left_join(ALL.ASVs, by = c("representative" = "Hash")) %>% 
 # group_by(representative) %>% 
  summarise(tot = sum(nReads)) # 2.7M reads
clasif.hashes %>% 
  filter(rank == "" & !is.na(score)) %>% # 
  left_join(ALL.ASVs, by = c("representative" = "Hash")) %>% 
 # group_by(representative) %>% 
  summarise(tot = sum(nReads)) # 0
 clasif.hashes %>% 
   filter(rank !="") %>% #
   left_join(ALL.ASVs, by = c("representative" = "Hash")) %>% 
 mutate(Sample_name = as.character(Sample_name)) %>%
   group_by(Sample_name) %>% 
  summarise(tot = sum(nReads)) %>% 
   ggplot(aes(x=fct_reorder(Sample_name, tot), y = tot))+
   geom_col()# 
clasif.hashes %>% 
  filter(rank == "")
 
# Create the species- level dataset
map_dfr(l2, ~(dplyr::count(.,rank)), .id = "Threshold" ) -> Summary.thresholds.df# This puts the results in a df


```

summarize by the sum of reads per hash
```{r df of reads per threshold}
  ALL.ASVs %>% group_by(Hash) %>% 
  summarise(n2=sum(nReads)) -> reads.per.hash
map_dfr(l2, ~(dplyr::add_count(.,rank)), .id = "Threshold" ) -> Summary.thresholds.2# This puts the results in a df
```


create plots to visualize the differences in classification bewteen thresholds to see that everything is working
```{r create plots for thresholds}
Summary.thresholds.2 %>% left_join(reads.per.hash, by = c("representative"="Hash")) -> Summary.thresholds.2
Summary.thresholds.2 %>% group_by(Threshold, rank, n) %>% summarise(nReads = sum(n2)) -> Summary.thresholds
unique(Summary.thresholds$rank)
#Summary.thresholds$rank <- fct_relevel(Summary.thresholds.df$rank, "class",   "infraclass", "superorder" , "order" , "suborder", "family", "subfamily" , "tribe",  "genus", "species", "clade", "cohort", "no rank") #this section isn't working to actually re-order the ranks while keeping the reads together
Summary.thresholds %>%
  
  ggplot (aes (x= rank, y = n, fill = Threshold)) +
  geom_col(position = "dodge", color = "black") +
  theme(axis.text.x = element_text(angle = 45,vjust = 0.75)) -> p
  
Summary.thresholds %>%
  
  ggplot (aes (x= rank, y = nReads, fill = Threshold)) +
  geom_col(position = "dodge", color = "black") +
  theme(axis.text.x = element_text(angle = 45,vjust = 0.75)) -> q
  ggsave(filename = "Different.thresholds.png", width = 14)  
  p+q+plot_layout(guides = "collect")
  ggsave(filename = "Different.thresholds.compare.png", width = 14)
```

now we want to make a combined rds file with the previous effort and the new one then a tax.table for use in the phyloseq object
```{r using tax.table.R to make a tax table}
combined <- rbind(previous.effort,clasif.hashes)
saveRDS(combined, file=paste0(Sys.Date(),"hashes.annotated.rds"))
#need to make this line not contigent on local directory
source(here("functions", "tax.table.R"))

taxtable <- tax.table(combined)

#make this a save as rds and save it in the github output folder 
write.csv(taxtable,file=paste0(Sys.Date(),"taxtable.csv"))#save it in the local directory for any future needs
```

