---
title: "tax_asvs_to_phyloseq"
author: "Erin D'Agnese"
date: "8/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Need to take the hashes annotated .rds in the github outputs, use the ASV table with hashes and make it
into a phyloseq object


```{r libraries and functions}
library(tidyverse)
library(here)
library(phyloseq)

```
load source function scripts
```{r}
source(here("functions","asv.matrix.R"))
source(here("functions","making.phyloseq.obj.R"))
source(here("functions", "tax.table.R"))
```


load the taxonomy file from the project file
```{r}
tax.input <- read_rds(here("Output","insect_output_files", "12S_testrun", "hashes.annotated.rds"))
asv.mat <- read_csv(here("Output", "ASV_metadata", "test_run_20210527", "12S", "ASV_table.csv"))
metadata <- read_csv(here("Output", "ASV_metadata", "test_run_20210527", "20210528_testseq_creek_metadata.csv"))
```
ONLY USE THIS FOR THE 12S ASV TABLE IF THEY AREN'T SPLIT
```{r}
asv_fish <- asv.mat[asv.mat$Locus =="MiFish",]
asv_mam <- asv.mat[asv.mat$Locus =="MiMammal",]
#lets assign mifish to asv.mat for the future work first
asv.mat <- asv_fish
```

Select out the metadata for the locus of choice
```{r}
metadata <- metadata[metadata$locus == "MiFish",]
```


Making the ASV table into a matrix 
```{r ASV pivot}
asv.mat <- asv.matrix(asv.mat)
```

making annotated hashes into the taxtable
```{r}
#first reorder so that taxon is after species
tax.input <- tax.input %>% relocate(taxon, .after = species)
tax.mat <- tax.table(tax.input)
```
making a phyloseq object from the input
```{r}
phylo <- make.phyloseq.obj(asv.mat, metadata, tax.mat)
```

check that things are working with a barplot
```{r}
plot_bar(phylo, fill = "family")
```

filter out the bacteria and uncharacterized/no rank ASVs
```{r}
phylo <- subset_taxa(phylo, !is.na(kingdom) & !kingdom %in% c("","no rank", "uncharacterized"))
```

check again to make sure they were removed
```{r}
plot_bar(phylo, fill = "kingdom")
```


then collapse to taxon and produce an OTU table to be combined with other run OTU tables later
```{r}
taxon_glom <- tax_glom(phylo, taxrank = "taxon")
#check that it kept all the column data

```

