---
title: "tax_asvs_to_phyloseq"
author: "Erin D'Agnese"
date: "8/2/2021"
output: 
  html_notebook:
    toc: true
params:
  folder:
    value: C:\Users\erdag\github\NextGenNEPA\Output\phyloseq_output\run3\COI
  hashes.annotated:
    value: C:\Users\erdag\github\NextGenNEPA\Output\hashes_to_taxonomy_output\run3\MiX\2021-10-18taxtable.csv
  ASV: 
   value: C:\Users\erdag\github\NextGenNEPA\Output\dada2_output_files\run3_20211018\COI_fwds_ASV_table.csv
  metadata:
      value: C:\Users\erdag\github\NextGenNEPA\Input\sequencing_metadata_files\master_sequencing_datasheet20211018.csv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = params$folder)
```

Need to take the hashes annotated .rds in the github outputs, use the ASV table with hashes and make it
into a phyloseq object


```{r libraries and functions}
library(tidyverse)
library(here)
library(phyloseq)
```
load source function scripts
```{r}
source(here("functions","asv.matrix.R"))
source(here("functions","making.phyloseq.obj.R"))
source(here("functions", "tax.table.R"))
```


load the taxonomy file from the project file
```{r}
tax.input <- read_csv(params$hashes.annotated)
asv.mat <- read_csv(params$ASV)
metadata <- read_csv(params$metadata)
```
ONLY USE THIS FOR THE 12S ASV TABLE IF THEY AREN'T SPLIT
```{r}
asv_fish <- asv.mat[asv.mat$Locus =="MiFish",]
asv_mam <- asv.mat[asv.mat$Locus =="MiMammal",]
#lets assign mifish to asv.mat for the future work first
asv.mat <- asv_mam

```

Select out the metadata for the correct run and locus of choice
```{r}
metadata <- metadata[metadata$Sequencing.run == "3",]
#metadata <- metadata[metadata$Locus == "MiFish",]
metadata <- metadata[metadata$Locus == "MiMammal",]
#metadata <- metadata[metadata$Locus == "COI",]
```


Making the ASV table into a matrix 
```{r ASV pivot}
asv.mat <- asv.matrix(asv.mat)
```

making annotated hashes into the taxtable
```{r}
#first reorder so that taxon is after species
tax.input <- tax.input %>% relocate(taxon, .after = species)
tax.mat <- tax.table(tax.input)
```
making a phyloseq object from the input
```{r}
phylo <- make.phyloseq.obj(asv.mat, metadata, tax.mat)
```

check that things are working with a barplot
```{r}
plot_bar(phylo, fill = "phylum")
```

filter out the bacteria and uncharacterized/no rank ASVs
```{r}
phylo_clean <- subset_taxa(phylo, !is.na(kingdom) & !kingdom %in% c("","no rank", "uncharacterized"))
```

check again to make sure they were removed
```{r}
plot_bar(phylo_clean, fill = "family")
```


then collapse to taxon and produce an OTU table to be combined with other run OTU tables later
```{r}
taxon_glom <- tax_glom(phylo_clean, taxrank = "taxon")
#check that it kept all the column data

```


making the ASV table with taxonomy rather than hashes including taxa that are uncharacterised
```{r}
##make a temporary phyloseq object so we don't alter the full original one
tmp.phylo <- phylo
hash.id <- taxa_names(phylo)
# this changes the header from the actual sequence to Seq_001, Seq_002 etc
taxa_names(tmp.phylo)
n_seqs <- seq(ntaxa(tmp.phylo))
len_n_seqs <- nchar(max(n_seqs))
taxa_names(tmp.phylo) <- paste("Seq", formatC(n_seqs, 
                                            width = len_n_seqs, 
                                            flag = "0"), sep = "_")
taxa_names(tmp.phylo)

# generate a vector containing the full taxonomy path for all OTUs
wholetax <- do.call(paste, c(as.data.frame(tax_table(tmp.phylo))
                  [c("kingdom", "phylum", "class", "order", "family", "genus", "species")], 
                  sep = "__"))  # to distinguish from "_" within tax ranks

#export the ASV table into a dataframe
ASV1 <- as.data.frame(otu_table(tmp.phylo))

#generate the vector of new seq_numbers
tmp <- row.names(ASV1)

# paste wholetax and OTU_ids together
for(i in 1:length(tmp)){
names(tmp)[i] = paste(wholetax[i], tmp[i], sep = "__")
}
#generate a mapping dataframe with the Seq_number and hash ids to keep 
map.ASV <- data.frame(hash.id, tmp)
write.csv(map.ASV,file=paste0(Sys.Date(),".map.ASV.hash.to.taxa.csv"))

# overwrite old row names
row.names(ASV1) <- names(tmp)
head(ASV1)


saveRDS(ASV1, file=paste0(Sys.Date(),".ASV.with.taxa.rds"))
write.csv(ASV1, file=paste0(Sys.Date(),".ASV.with.taxa.csv"))
```

Making an agglomerated OTU table with counts by taxon as above
```{r}
##make a temporary phyloseq object so we don't alter the full original one
tmp.tg <- taxon_glom
tmp.hash <- taxa_names(taxon_glom)
# this changes the header from the actual sequence to Seq_001, Seq_002 etc
taxa_names(tmp.tg)
n_seqs <- seq(ntaxa(tmp.tg))
len_n_seqs <- nchar(max(n_seqs))
taxa_names(tmp.tg) <- paste("Taxon", formatC(n_seqs, 
                                            width = len_n_seqs, 
                                            flag = "0"), sep = "_")

# generate a vector containing the full taxonomy path for all OTUs
wholetax.tg <- do.call(paste, c(as.data.frame(tax_table(tmp.tg))
                  [c("kingdom", "phylum", "class", "order", "family", "genus", "species")], 
                  sep = "__"))  # to distinguish from "_" within tax ranks


#export the taxon agglomerated OTU table to a dataframe 
OTU1 <- as.data.frame(otu_table(tmp.tg))

#generate a vector of the taxon ids
tmp.tx <- row.names(OTU1)

#generate a mapping dataframe with the Seq_number and hash ids to keep 
map.taxon <- data.frame(tmp.hash, tmp.tx, wholetax.tg)
write.csv(map.taxon,file=paste0(Sys.Date(),".map.taxon.glommed.hash.csv") )

# paste wholetax and Taxon_numbers together
for(i in 1:length(tmp.tx)){
names(tmp.tx)[i] = paste(wholetax.tg[i], tmp.tx[i], sep = "__")
}
# overwrite old row names
row.names(OTU1) <- names(tmp.tx)
head(OTU1)

saveRDS(OTU1, file=paste0(Sys.Date(),".OTU.taxon.with.taxa.rds"))
write.csv(OTU1, file=paste0(Sys.Date(),".OTU.taxonASV.with.taxa.csv"))
```


