---
title: "working_ANCOM"
author: "Erin D'Agnese"
date: "6/28/2021"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
coi.otu <- read.csv("C:/Users/erdag/Documents/NextgenNEPA/analysis/test_run_20210527/LerayXT.besthit.OTUtable.csv", header=T, check.names=FALSE)
row.names(coi.otu) <- coi.otu[,1]#rename row.names to the hash id
coi.otu <- coi.otu[,-1]#remove the first column which is now the row.name
coi.otu <- coi.otu[,-43]#remove the kangaroo sample

```

run asv.matrix function to get the ASV tables into the format necessary and remove the kangaroo
```{r}
setwd("C:/Users/erdag/Documents/NextgenNEPA/analysis")
source("asv.matrix.R")
coi.asv.mat <- as.matrix(coi.asv)
coi.asv.mat <- coi.asv.mat[,-43] #remove the column of Roo sequences
```


get the metadata file - use the metadata file without roo
```{r}
metadata <-read.csv("C:/Users/erdag/Documents/NextgenNEPA/analysis/20210528_testseq_creek_metadata_noRoo.csv",  row.names=1, header=T)
metadata <- metadata[metadata$locus == "LerayXT",]
```


```{r}
source("ancom_v2.1.R")
```

create the feature table for ANCOM

```{r}
feature_table = coi.otu; sample_var = "sample_id"; group_var = NULL
```

preprocessing of feature table and metadata table to remove outliers
```{r}
out_cut = 0.05; zero_cut = 0.98; lib_cut = 0; neg_lb = TRUE
prepro = feature_table_pre_process(feature_table, meta_data, sample_var, group_var,
                                   out_cut, zero_cut, lib_cut, neg_lb)
feature_table = prepro$feature_table # Preprocessed feature table
meta_data = prepro$meta_data # Preprocessed metadata
struc_zero = prepro$structure_zeros # Structural zero info
```


run ancom for creek level differential abundance while accounting for random
```{r}
main_var = "Event"; p_adj_method = "BH"; alpha = 0.05
adj_formula = NULL ; rand_formula = NULL
t_start = Sys.time()
res = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method,
            alpha, adj_formula, rand_formula)
t_end = Sys.time()
t_run = t_end - t_start

write_csv(res$out, "LerayXT_Event_ANCOM.csv")
```

visualise the volcano plot of the output

```{r}
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")

# Annotation data
dat_ann = data.frame(x = min(res$fig$data$x), y = cut_off["detected_0.6"], label = "W[0.6]")

fig = res$fig +
  geom_hline(yintercept = cut_off["detected_0.6"], linetype = "dashed") +
  geom_text(data = dat_ann, aes(x = x, y = y, label = label),
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig
```

```{r}
df1 <- res$out
df2 <- sum.table.event
df3 <- merge(df1,df2, by=c('taxa_id'),all.x=TRUE, all.y=TRUE)
write.csv(df3, "C:/Users/erdag/Documents/NextgenNEPA/analysis/test_run_20210527/LerayBesthit.even.ANCOM.sums.mapped.csv", row.names = FALSE)
df4 <- read.csv("C:/Users/erdag/Documents/NextgenNEPA/analysis/test_run_20210527/LerayXT.taxon.TAXtable.csv", header = T)
df4$taxa_id <- df4$X
df4 <- df4[,-1]
df5 <- merge(df3,df4, by=c('taxa_id'), all.x=TRUE, all.y=TRUE)
write.csv(df5, "C:/Users/erdag/Documents/NextgenNEPA/analysis/test_run_20210527/LerayBesthit.event.ANCOM.sums.mapped.withtaxa.csv", row.names = FALSE)
```


by creek when accounting for month as a covariate
```{r}
main_var = "Creek"; p_adj_method = "BH"; alpha = 0.05
adj_formula = "Month" ; rand_formula = NULL
t_start = Sys.time()
res = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method,
            alpha, adj_formula, rand_formula)
t_end = Sys.time()
t_run = t_end - t_start

write_csv(res$out, "LerayXT_CreekbyMonth_ANCOM.csv")
```

visualise the volcano plot of the output

```{r}
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")

# Annotation data
dat_ann = data.frame(x = min(res$fig$data$x), y = cut_off["detected_0.6"], label = "W[0.6]")

fig = res$fig +
  geom_hline(yintercept = cut_off["detected_0.6"], linetype = "dashed") +
  geom_text(data = dat_ann, aes(x = x, y = y, label = label),
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig
```

```{r}
df1 <- res$out
df2 <- sum.table.creek
df3 <- merge(df1,df2, by=c('taxa_id'),all.x=TRUE, all.y=TRUE)
write.csv(df3, "C:/Users/erdag/Documents/NextgenNEPA/analysis/test_run_20210527/LerayBesthit.CreekbyMonth.ANCOM.sums.mapped.csv", row.names = FALSE)
df4 <- read.csv("C:/Users/erdag/Documents/NextgenNEPA/analysis/test_run_20210527/LerayXT.taxon.TAXtable.csv", header = T)
df4$taxa_id <- df4$X
df4 <- df4[,-1]
df5 <- merge(df3,df4, by=c('taxa_id'), all.x=TRUE, all.y=TRUE)
write.csv(df5, "C:/Users/erdag/Documents/NextgenNEPA/analysis/test_run_20210527/LerayBesthit.CreekbyMonth.ANCOM.sums.mapped.withtaxa.csv", row.names = FALSE)
```

by station with creek as covariate
```{r}
main_var = "Station"; p_adj_method = "BH"; alpha = 0.05
adj_formula = "Creek" ; rand_formula = NULL
t_start = Sys.time()
res = ANCOM(feature_table, meta_data, struc_zero, main_var, p_adj_method,
            alpha, adj_formula, rand_formula)
t_end = Sys.time()
t_run = t_end - t_start

write_csv(res$out, "LerayXT_StationbyCreek_ANCOM.csv")
```

visualise the volcano plot of the output

```{r}
n_taxa = ifelse(is.null(struc_zero), nrow(feature_table), sum(apply(struc_zero, 1, sum) == 0))
# Cutoff values for declaring differentially abundant taxa
cut_off = c(0.9 * (n_taxa -1), 0.8 * (n_taxa -1), 0.7 * (n_taxa -1), 0.6 * (n_taxa -1))
names(cut_off) = c("detected_0.9", "detected_0.8", "detected_0.7", "detected_0.6")

# Annotation data
dat_ann = data.frame(x = min(res$fig$data$x), y = cut_off["detected_0.6"], label = "W[0.6]")

fig = res$fig +
  geom_hline(yintercept = cut_off["detected_0.6"], linetype = "dashed") +
  geom_text(data = dat_ann, aes(x = x, y = y, label = label),
            size = 4, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE)
fig
```

```{r}
df1 <- res$out
df2 <- sum.table.event
df3 <- merge(df1,df2, by=c('taxa_id'),all.x=TRUE, all.y=TRUE)
write.csv(df3, "C:/Users/erdag/Documents/NextgenNEPA/analysis/test_run_20210527/LerayBesthit.CreekbyStation.ANCOM.sums.mapped.csv", row.names = FALSE)
df4 <- read.csv("C:/Users/erdag/Documents/NextgenNEPA/analysis/test_run_20210527/LerayXT.taxon.TAXtable.csv", header = T)
df4$taxa_id <- df4$X
df4 <- df4[,-1]
df5 <- merge(df3,df4, by=c('taxa_id'), all.x=TRUE, all.y=TRUE)
write.csv(df5, "C:/Users/erdag/Documents/NextgenNEPA/analysis/test_run_20210527/LerayBesthit.CreekbyStation.ANCOM.sums.mapped.withtaxa.csv", row.names = FALSE)
```
